### 2024-02-09 ms.qmd
Updates

Our results suggest that the early environment manipulation did not affect learning; we found that the estimated learning slopes were not different between treatments or species. However, learning slopes were lower when the blue feeders were the correct choice compared to those lizards assigned to the group 'Red' for _L. delicata_, but not for _L. guichenoti_. This difference could be due to a potential bias towards blue in the initial choice, as we found that the estimated probability of choosing the correct feeder in the first trial was higher than expected by chance when the correct feeder was blue (see Supplementary Material). Moreover, those treatments and species whose learning slopes do not differ significantly from zero (see Table 1), are associated to highest bias (see Supplementary Material). Because there was a possibility that the bias towards the blue feeder was something acquired during training, we compared the light spectrum of the ramps used in the associative task and the white ones used during habituation, but we did no see any relevant similarity between the white and blue ramps (see Supplementary Material). As such, we consider this bias to not be a derivate from our experimental design. The species confidence hypothesis posits that animals should bias their attention toward colours they are familiar with, and empirical studies show that, for instance, dark and light blue T-shirts were associated with lower flight initiation distances and higher capture rates in Western fence lizards where blue is used in intraspecific communication [@putman2017fear]. However, to the best of our knowledge, noone has reported a colour bias in _L. delicata_ or _L. guichenoti_ before; and it is not considered to be involved in intraspecific communication [@torr1996patterns, @chapple_biology_2014]. Still, we recommend to select carefully the colours employed in associative tasks with these species, especially _L. delicata_, for future cognitive tasks employing visual stimuli. Nevertheless, when lizards assigned to blue or red were analysed separately, the between treatments comparissons remain the same in both species. Furthermore, the non-significant trends we observed were not consistent between species or colour assigned. As such, we conclude that, overall, lizards learned at the same rate regardless of the early environment manipulation.  

@@

>>>>

Our results suggest that the early environment manipulation did not affect learning; we found that the estimated learning slopes were not different between treatments or species. However, learning slopes were lower when the blue feeders were the correct choice compared to those lizards assigned to the group 'Red' for _L. delicata_, but not for _L. guichenoti_. This difference could be due to a potential bias towards blue in the initial choice, as we found that the estimated probability of choosing the correct feeder in the first trial was higher than expected by chance when the correct feeder was blue (see Supplementary Material). Moreover, those treatments and species whose learning slopes do not differ significantly from zero (see Table 1), are associated to highest bias (see Supplementary Material). Because there was a possibility that the bias towards the blue feeder was something acquired during training, we compared the light spectrum and the perceived chromatic differences between the ramps used in the associative task and the white ones used during habituation, but we did no see any relevant similarity between the white and blue ramps (see Supplementary Material). As such, we consider this bias to not be a derivate from our experimental design. The species confidence hypothesis posits that animals should bias their attention toward colours they are familiar with, and empirical studies show that, for instance, dark and light blue T-shirts were associated with lower flight initiation distances and higher capture rates in Western fence lizards where blue is used in intraspecific communication [@putman2017fear]. However, to the best of our knowledge, noone has reported a colour bias in _L. delicata_ or _L. guichenoti_ before; and it is not considered to be involved in intraspecific communication [@torr1996patterns, @chapple_biology_2014]. Still, we recommend to select carefully the colours employed in associative tasks with these species, especially _L. delicata_, for future cognitive tasks employing visual stimuli. Nevertheless, when lizards assigned to blue or red were analysed separately, the between treatments comparissons remain the same in both species. Furthermore, the non-significant trends we observed were not consistent between species or colour assigned. As such, we conclude that, overall, lizards learned at the same rate regardless of the early environment manipulation.  

@@

#### Light spectrum
To test if the bias towards blue was something acquired during training, we compared the light spectrum of the ramps used in the associative task and the white ones used during habituation. We took three measurements of ten feeders per colour with a spectophotometer, and then analysed the spectrum and the perceived differences in colour using the package pavo. The spetrum of each type of feeder are shown in @fig-spectrum; the perceived chromatic constrasts between feeders are shown in @fig-perceived1 and @fig-perceived2.  
```{r, fig-spectrum}
#| label: fig-spectrum
#| fig.cap: "Light spectrum of the ramps used in the associative task and the white ones used during habituation. The different colours represent the different ramps."
#
# Load required packages
pacman::p_load(pavo,plotrix,plyr,MCMCglmm,Hmisc,DIZutils)
#
# For importing the files, change the extension of the .jaz files to .txt files!  
# Import spec data for pablos feeders - CHNAGE TO WHERE YOUR FILES ARE LOCATED
specs <- getspec("./Others/Feeders", lim = c(300,700), ext = 'txt') %>%
  procspec(specs, opt = "smooth" , span = 0.2, fixneg = "addmin") 
# Plot smoothed raw data as reflectance curve with approximate rgb colours
plot(specs, type = "o", col = spec2rgb(specs))
```

```{r, fig-perceived1}
#| label: fig-perceived1
#| fig.cap: "Perceived chromatic contrasts between feeders."
#
## Processing the data with the lizard visual model 
# Processing using lizard vision model provided for Dan and Martin's Phyrnocephalus mystaceus paper
liz_vis <- sensmodel(c(360, 440, 493, 571)) 
#
# Apply this over the feeder data visual as the visual system, achromatic is the receptor sensitivity, illum is the illuminate (here using an ideal white illuminate preset), bkg is the background against which we are measuring (here using an ideal background with no influence). Other conditions are applied as used in the P mystaceus paper.
# Does produce an error about luminance but this is noted on the original P mystaceus paper code and should be okay
# summary(feeders_lizardview) can be used to check what conditions were applied in the visual model
feeders_lizardview <- vismodel(specs, visual = liz_vis, achromatic = "l", illum = "ideal", bkg = "ideal", vonkries = TRUE,   relative = FALSE, qcatch = "fi")
#                             
# Calculate colour distances between all feeder measurements
coldistance_feeders <- coldist(feeders_lizardview,achro = TRUE, noise = "neural", n = c(1, 1, 3.5, 6),  weber = 0.1)
#
# Name all contrasts e.g white/green to categorise them
annotated_coldistance_feeders <- coldistance_feeders %>%
  separate(patch1, into = c("color1", "number1"), sep = "_") %>%
  separate(patch2, into = c("color2", "number2"), sep = "_") %>%
  mutate(color_comparison = paste0(color1, "/", color2)) %>%
  mutate(color_comparison = paste0(pmin(color1, color2), "/", pmax(color1, color2))) %>% # Remove duplicate categories(e.g white/green is the same as green/white and all are called green/white)
  subset(!(color_comparison %in% c("B/B", "G/G", "R/R", "W/W")))
#
# Plot the colour differences between the feeders: 
# Note - above 1 is expected to be visually different under the selected visual system (here, lizard view)
# dS, or chromatic contrast
dScomparisons <- ggplot(annotated_coldistance_feeders, aes(x = color_comparison, y = dS, fill = color_comparison)) + 
  geom_boxplot() + 
  theme_classic() +
  ylab("Chromatic Contrast (dS)") +
  xlab("Feeder Colours Comparison") +
 scale_fill_manual(values =c("B/G" = 'steelblue',  "B/R" = 'purple', "B/W" = 'lightblue', "G/R" = 'brown', "G/W" = 'lightgreen', "R/W" = 'red')) +
 theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) + 
  theme(
    axis.title = element_text(size = 15, family = "Times New Roman"),
    axis.text = element_text(size = 10, family = "Times New Roman"),
    legend.title = element_text(size = 12, family = "Times New Roman"),
    legend.text = element_text(size = 10, family = "Times New Roman"),
    strip.text = element_text(size = 13, family = "Times New Roman")
  ) 
print(dScomparisons)
```

```{r, fig-perceived2}
#| label: fig-perceived2
#| fig.cap: "Perceived achromatic contrasts between feeders."
#dL, or achromatic contrast
dLcomparisons <- ggplot(annotated_coldistance_feeders, aes(x = color_comparison, y = dL, fill = color_comparison)) + 
  geom_boxplot() + 
  theme_classic() +
  ylab("Achromatic Contrast (dL") +
  xlab("Feeder Colours Comparison") +
 scale_fill_manual(values =c("B/G" = 'steelblue',  "B/R" = 'purple', "B/W" = 'lightblue', "G/R" = 'brown', "G/W" = 'lightgreen', "R/W" = 'red')) +
 theme(plot.margin = margin(5.5, 5.5, 5.5, 5.5, "mm")) + 
  theme(
    axis.title = element_text(size = 15, family = "Times New Roman"),
    axis.text = element_text(size = 10, family = "Times New Roman"),
    legend.title = element_text(size = 12, family = "Times New Roman"),
    legend.text = element_text(size = 10, family = "Times New Roman"),
    strip.text = element_text(size = 13, family = "Times New Roman")
  )
print(dLcomparisons)
```
